---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

pacman::p_load(GeoLink, sf)

```

# GeoLink

<!-- badges: start -->
[![R-CMD-check](https://github.com/SSA-Statistical-Team-Projects/GeoLink/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/SSA-Statistical-Team-Projects/GeoLink/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of GeoLink is to support the download of high resolution geospatial/remote sensing data, zonal statistics computation, shapefile tesselation as well as spatial joining of geospatial data with unit level georeferenced data (such as household surveys).

## Installation

You can install the development version of GeoLink from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("SSA-Statistical-Team-Projects/GeoLink")
```

## Usage

There are several use cases intended for the package. The process begins with downloading geospatial data (typically raster format data) and: 

- Estimating geospatial indicator value within a certain area of the unit level data. For instance, estimating the level of economic activity around each household location within a household survey by using the night time lights extraction.  

```{r, message = FALSE}

## estimate night time luminosity (NTL) within a 100m of each household location in the year 2020

geolink_ntl(start_date = "2020-01-01",
            end_date = "2020-12-31",
            annual_version = "v21",
            indicator = "average_masked",
            survey_dt = st_as_sf(hhgeo_dt[ADM1_EN == "Abia",],
                                 crs = 4326),
            buffer_size = 100,
            shp_dt = shp_dt[shp_dt$ADM1_EN == "Abia",])



```


- Estimating a spatial indicator within the grid tesselation that a household is location wtihin. This is often useful for those interested in poverty mapping or other model building exercises that require equal area grids (option available within the function as `grid_size` argument) for an entire area
```{r, message = FALSE}

## estimate night time luminosity (NTL) for household location based on the tesselation of a shapefile i.e. for the grid in which the household is located in the year 2020


geolink_ntl(start_date = "2020-01-01",
            end_date = "2020-12-31",
            annual_version = "v21",
            indicator = "average_masked",
            survey_dt = st_as_sf(hhgeo_dt[ADM1_EN == "Abia",],
                                 crs = 4326),
            grid_size = 1000,
            shp_dt = shp_dt[shp_dt$ADM1_EN == "Abia",])



```


- Simply linking the zonal statistics to a shapefile (tesselated or simply at native resolution)
```{r, message = FALSE}

## estimate night time luminosity (NTL) for a tesselated shapefile

geolink_ntl(start_date = "2020-01-01",
            end_date = "2020-12-31",
            annual_version = "v21",
            indicator = "average_masked",
            grid_size = 1000,
            shp_dt = shp_dt[shp_dt$ADM1_EN == "Abia",])



## estimate NTL for the shapefile

geolink_ntl(start_date = "2020-01-01",
            end_date = "2020-12-31",
            annual_version = "v21",
            indicator = "average_masked",
            grid_size = NULL,
            shp_dt = shp_dt[shp_dt$ADM1_EN == "Abia",])



```



























